---
title: "Data Manipulation"
author: "Ashley Tseng"
date: "9/19/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(dplyr)
```

## Import datasets

```{r}
litters_data = read_csv("./data/FAS_litters.csv")
litters_data = janitor::clean_names(litters_data)

pups_data = read_csv("./data/FAS_pups.csv")
pups_data = janitor::clean_names(pups_data)

```
# R doesn't know about the "group" variable inside of the litters dataset "litters_data". We need to tell R where to look: "pull(litters_data, group).



## Selecting

# If we only want to select certain columns (variables) from the dataset "litters_data", then use select():
```{r}
select(litters_data, group, litter_number)

select(litters_data, group, litter_number, gd0_weight)

# Can use a helper function like "starts_with" to tell R to select all columns from the dataset that start with "pups":
select(litters_data, group, litter_number, gd0_weight, starts_with("pups"))

select(litters_data, litter_number, group, gd0_weight)

# Tells R that we want litters_number, then group, then everything else to come after:
select(litters_data, litter_number, group, everything())

# Tells R that we want all the columns in the dataset selected EXCEPT for group (denoted by the -):
select(litters_data, -group)

# Keeps litter_number, then extract every column from my dataset between gd0_weight and pups_born_alive:
select(litters_data, litter_number, gd0_weight:pups_born_alive)

# We can rename column names with the select function too (but again, this isn't permanent):
select(litters_data, GROUP = group, litter_number)

# We can also rename column names with the rename() function (this also is not permanent!):
rename(litters_data, GROUP = group)

```
# Note that with selecting, we aren't making any changes to the existing dataframe unless we make a new df. 



## Filtering

# If we only want to select certain columns (variables) from the dataset "litters_data", then use select():

```{r}
filter(litters_data, group == "Con7")

filter(litters_data, gd_of_birth == 20)

filter(litters_data, gd_of_birth < 20)

filter(litters_data, gd_of_birth >= 20)

filter(litters_data, pups_born_alive < 6)

filter(litters_data, pups_born_alive < 6, group == "Con7")

# We want to filter for pups that survive that aren't equal to 4:
filter(litters_data, pups_survive != 4)

# Can use not and have the function inside the ():
# !(pups_survive == 4)

# If we want to filter for "Con7" or "Mod8":
filter(litters_data, group == "Con7" | group == "Mod8")
# This is equivalent to the or function above, but Prof. prefers this one: 
filter(litters_data, group %in% c("Con7", "Mod8"))

filter(litters_data, pups_born_alive >= 4, pups_born_alive <= 6)

# If we want to filter for groups where the sum of gd0_weight and gd18_weight is less than 70:
filter(litters_data, gd0_weight + gd18_weight < 70)

# DO NOT DO THIS:
filter(litters_data, !is.na(gd0_weight))

# Instead, do this (tells R that we want to drop the missing values for "gd0_weight" in the "litters_data" dataframe:
drop_na(litters_data, gd0_weight)

```



## Mutate

# Sometimes you need to select columns; sometimes you need to change them or create new ones. You can do this using mutate:
```{r}
mutate(
  litters_data,
  wt_gain = gd18_weight - gd0_weight,
  group = str_to_lower(group)
)

# "wt_gain" variable is a new variable created from old variables
# We modified the first letter of the "group" variable names from upper case to lower case
```



## Arrange

```{r}
arrange(litters_data, group, pups_born_alive)
# R has put our dataset in order based on the number of pups born alive

arrange(litters_data, group, desc(pups_born_alive))
# R has put our dataset in descending order based on the number of pups born alive

arrange(litters_data, pups_born_alive, gd0_weight)
# Within the litters of pups born alive, the groups are also sorted by their gd0_weight

head(arrange(litters_data, group, pups_born_alive), 10)
```



