---
title: "Data Import"
author: "Ashley Tseng"
date: "9/17/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(readxl)
library(haven)
```

## Load in the litters dataset

```{r}
## read in a dataset
litters_data = read_csv(file = "./data/FAS_litters.csv")
## using a relative path "./data"
## don't ever use "read.csv" to read in a dataset, always use "read_csv"
## the first thing you should do after reading in a dataset is print the dataframe; can also use View(litters_data)

## rewrite to clean the column names
litters_data = janitor::clean_names(litters_data)
```

## Load in the pups dataset

```{r}
pups_data = read_csv(file = "./data/FAS_pups.csv")
pups_data = janitor::clean_names(pups_data)
```

## To skim summary statistics
skimr::skim(litters_data)

## If we wanted to skip the first 50 lines of data and not assume the first row are variable names:
```{r}
litters_data = read_csv(file = "./data/FAS_litters.csv",
  skip = 10, col_names = FALSE)
## Parsed: trying to guess what variable type each variable is
## What read_csv does it try to guess what variable type each type is based on skimming the data
## "double" is computer jargon for number
```

## Parsing columns
```{r}
litters_data = read_csv(file = "./data/FAS_litters.csv",
  col_types = cols(
    Group = col_character(),
    `Litter Number` = col_character(),
    `GD0 weight` = col_double(),
    `GD18 weight` = col_double(),
    `GD of Birth` = col_integer(),
    `Pups born alive` = col_integer(),
    `Pups dead @ birth` = col_integer(),
    `Pups survive` = col_integer()
  )
)

## We see that we get the same column format when parsing and when not parsing
## In R, valid column names cannot include spaces. For columns that do include spaces, need to use back ticks ``
```

## Reading in an Excel file
```{r}
mlb11_data = read_excel("./data/mlb11.xlsx", n_max = 20)
head(mlb11_data, 5)

## This works too
mlb11_data_2 = read_excel("data/mlb11.xlsx", n_max = 20)

## This doesn't work
mlb11_data_2 = read_excel("mlb11.xlsx", n_max = 20)
```

## Reading in a SAS file
```{r}
pulse_data = read_sas("./data/public_pulse_data.sas7bdat")
```

## To pull a variable from a data frame:
pull(data_frame, variable_name)
## read_csv is bringing in data and making it a tibble; read.csv is making it a data frame

## If we try to use read.csv:
litters_data = read.csv(file = "./data/FAS_litters.csv")
class(pull(litters_data, Group))
## We get an error that object 'Group' not found

## Need to use read_csv:
litters_data = read_csv(file = "./data/FAS_litters.csv")
class(pull(litters_data, Group))
## "character" variable

```{r}
mlb11_data_subset = read_excel(
  path = "./data/mlb11.xlsx",
  range = "A1:D7"
)

## to export a .csv file
write_csv(mlb11_data_subset, path = "./data/mlb_subset.csv")
```