---
title: "Tidy Data"
author: "Ashley Tseng"
date: "9/24/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(dplyr)
```

## Wide to long

```{r}
pulse_data = 
  haven::read_sas("./data/public_pulse_data.sas7bdat") %>%
  janitor::clean_names()

pulse_data

pivot_longer(
  pulse_data,
  bdi_score_bl:bdi_score_12m,
  names_to = "visit",
  values_to = "bdi"
)
```
"names_to": The columns that I currently have are going to become their own variable "visit".
"values_to": Then take the values from those columns and put them their new column.
"pivot_longer" results in two new columns; wide to long format.


We didn't like that pivot_longer was a separate action from the haven and janitor action; let's pipe it all together: 
```{r}
pulse_data = 
  haven::read_sas("./data/public_pulse_data.sas7bdat") %>%
  janitor::clean_names() %>%
  pivot_longer(
  bdi_score_bl:bdi_score_12m,
  names_to = "visit",
  names_prefix = "bdi_score_",
  values_to = "bdi"
) %>% 
  mutate(
    visit = recode(visit, "bl" = "00m")
  )
```

## Separate in litters

```{r}
litters_data = 
  read_csv("./data/FAS_litters.csv", col_types = "ccddiiii") %>% 
  janitor::clean_names() %>%
  separate(group, into = c("dose", "day_of_tx"), sep = 3) %>%
  mutate(
    dose = str_to_lower(dose),
    wt_gain = gd18_weight - gd0_weight) %>%
  arrange(litter_number)

litters_data
```
"separate" is useful for separating one column into two columns. Where separate gets tricky is telling R WHERE to split something --> we will use "sep = character # where we want to separate at". This creates two new columns in the order you told R at the charcter # indicated.


## Go untidy. We can deliberately UNtidy something using "pivot_wider":

```{r}
analysis_result = tibble(
  group = c("treatment", "treatment", "placebo", "placebo"),
  time = c("pre", "post", "pre", "post"),
  mean = c(4, 8, 3.5, 4)
) %>% view

pivot_wider(  
  analysis_result, 
  names_from = "time", 
  values_from = "mean") %>% view
```


## 

```{r}

```


