---
title: "Tidy Data"
author: "Ashley Tseng"
date: "9/24/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(dplyr)
```

## Wide to long

```{r}
pulse_data = 
  haven::read_sas("./data/public_pulse_data.sas7bdat") %>%
  janitor::clean_names()

pulse_data

pivot_longer(
  pulse_data,
  bdi_score_bl:bdi_score_12m,
  names_to = "visit",
  values_to = "bdi"
)
```
"names_to": The columns that I currently have are going to become their own variable "visit".
"values_to": Then take the values from those columns and put them their new column.
"pivot_longer" results in two new columns; wide to long format.


We didn't like that pivot_longer was a separate action from the haven and janitor action; let's pipe it all together: 
```{r}
pulse_data = 
  haven::read_sas("./data/public_pulse_data.sas7bdat") %>%
  janitor::clean_names() %>%
  pivot_longer(
  bdi_score_bl:bdi_score_12m,
  names_to = "visit",
  names_prefix = "bdi_score_",
  values_to = "bdi"
) %>% 
  mutate(
    visit = recode(visit, "bl" = "00m")
  )
```

## Separate in litters

```{r}
litters_data = 
  read_csv("./data/FAS_litters.csv", col_types = "ccddiiii") %>% 
  janitor::clean_names() %>%
  separate(group, into = c("dose", "day_of_tx"), sep = 3) %>%
  mutate(
    dose = str_to_lower(dose),
    wt_gain = gd18_weight - gd0_weight) %>%
  arrange(litter_number)

litters_data
```
"separate" is useful for separating one column into two columns. Where separate gets tricky is telling R WHERE to split something --> we will use "sep = character # where we want to separate at". This creates two new columns in the order you told R at the charcter # indicated.


## Go untidy. We can deliberately UNtidy something using "pivot_wider":

```{r}
analysis_result = tibble(
  group = c("treatment", "treatment", "placebo", "placebo"),
  time = c("pre", "post", "pre", "post"),
  mean = c(4, 8, 3.5, 4)
) %>% view

pivot_wider(  
  analysis_result, 
  names_from = "time", 
  values_from = "mean") %>% view
```


## Binding rows. We want to first load the individual tables from the same Excel spreadsheet "LotR_Words.xlsx," then bind them into one dataframe.

```{r}
fellowship_ring = 
  readxl::read_excel("./data/LotR_Words.xlsx", range = "B3:D6") %>%
  mutate(movie = "fellowship_ring") %>% view

two_towers = 
  readxl::read_excel("./data/LotR_Words.xlsx", range = "F3:H6") %>%
  mutate(movie = "two_towers") %>% view

return_king = 
  readxl::read_excel("./data/LotR_Words.xlsx", range = "J3:L6") %>%
  mutate(movie = "return_king") %>% view

lotr_tidy = 
  bind_rows(fellowship_ring, two_towers, return_king) %>%
  janitor::clean_names() %>%
  pivot_longer(
    female:male,
    names_to = "sex", 
    values_to = "words") %>%
  mutate(race = str_to_lower(race)) %>% 
  select(movie, race, sex, words) %>% view

```


## Joining datasets. We want to first read in our datasets. Recoding the sex variable from numerical to their categories in the pups dataset.

```{r}
pup_data = 
  read_csv("./data/FAS_pups.csv", col_types = "ciiiii") %>% 
  janitor::clean_names() %>%
  mutate(sex = recode(sex, `1` = "male", `2` = "female")) %>% view

litter_data = 
  read_csv("./data/FAS_litters.csv", col_types = "ccddiiii") %>%
  janitor::clean_names() %>%
  select(-pups_survive) %>%
  mutate(
    wt_gain = gd18_weight - gd0_weight,
    group = str_to_lower(group))
```

Try to join these datasets!

We think that the pup dataset has all the variables that I need, so I want to pull everything from litters to the pup dataset. Take everything that I see in the pup dataset, match it to its corresponding litter number in the litter data, and add it to the pup dataset.
```{r}
fas_data = 
  left_join(pup_data, litter_data, by = "litter_number") %>% view
```
The first two pups are from the same litter number so we see that all the columns pulled from "litter_data" are exactly the same. This is exactly what we expect, so this is good!

We could hypothetically join by more than one column name:
fas_data = 
  left_join(pup_data, litter_data, by = c("litter_number", "something_else")


A left join and full join will NOT give you the same number of rows (we get two extra):
```{r}
full_join(pup_data, litter_data, by = "litter_number")
```


